---
title: "You should add a descriptive title here"
author: "Your Name Here"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Overview

\*\*\*\*\* FOR YOUR INFO - THIS TEXT CAN BE DELETED WHEN YOU START YOUR OWN SCRIPT \*\*\*\*\* This script is intended as a template to give a starting point for well-structured R scripts for DERT projects. The actual content will vary widely depending on the project, but code should we well organized (use headers and labels to break up sections, allow code chunks to be run easily, and enable easy navigation) and well documented (use comments to explain what you are doing and why). Note that the headings will show up in the navigation pane to the right (if you have "Outline" enabled), allowing you to easily jump to different sections of the script. The labels on the code chunks also allow you to easily find specific code chunks using the "Find in Document" feature (Ctrl+F or Cmd+F) or the "Overview" dropbox in the bottom left of the script window. The script uses a config file to define the path to the DERT Data & Metadata Dropbox folder on your computer. This allows you to read in and write out data and figures without hardcoding paths that will not work on other people's computers. \*\*\*\*\*

This script uses data ... description of the project, data source, etc ... to ... description of what it does ...

File Created:\
yyyy-mm-dd your name here

Revisions:\
yyyy-mm-dd your name and description of major revisions here ... add lines as needed

## Variable Names and Descriptions

Below are the variable names and meta-data for all variables in data that are imported or calculated in this script. \*\*\*\* UPDATE THE VARIABLE NAMES AS NEEDED FOR YOUR SCRIPT\*\*\*\*

#### Experimental design variables

`LitterbagID`: The unique identifier for each litterbag used in the experiment. Allowable values = M001 to M102 for mopane samples and D001 to D102 for dung samples.

`LitterType`: The type of litter being analyzed. Allowable values = mopane and dung.

`Block`: The block assigned to the litterbag. Litterbags were deployed in blocks of co-located litterbags (see "Ongava litterbag maps.pptx" for details). Allowable values = A - F.

`Microsite`: The microsite assigned to the litterbag. Allowable values = tree, open, NA. Tree microsites were under mopane canopies and open microsites were in open areas without tree canopy cover. Time 0 litterbags were not assigned to microsites and are indicated as NA.

`WaterTrt`:The water treatment assigned to the block that the litterbag was in. Allowable values = Ctrl, Addition, and NA. Time 0 litterbags were not assigned to a water treatment and are indicated as NA.

`Location`: The relative location within a block assigned to the litterbag to help in locating the bags for future collections (see "Ongava litterbag maps.pptx" for details). Allowable values = 1 - 8, NA. Time 0 litterbags were not assigned to a water treatment and are indicated as NA.

`CollectionTime`: The time point assigned for the litterbag to be collected. Allowable values = 0 - 6, with each 1 unit corresponding to 3 months.

`CollectionDate`: The date the litterbag was collected from the field. Format = YYYY-MM-DD.

`LitterbagStart`: The first litterbag number assigned to a particular WetDry correction sample. Allowable values = 1 - 102 for both dung and mopane.

`LitterbagEnd`: The last litterbag number assigned to a particular WetDry correction sample. Allowable values = 1 - 102 for both dung and mopane.

`Crucible`: The crucible number assigned to a particular ash mass sample. Allowable values = 1 - 15 for both dung and mopane.

`AshDate`: The date the ash mass sample was processed. Format = YYYY-MM-DD AM/PM.

#### Measured variables

`LitterbagMass`: The mass of the litterbag at the start of the experiment. Allowable values = 2.5-4 for dung and NA for mopane. Units = g.

`Litterbag_Litter`: The mass of the litterbag and the wet litter within the litterbag at the start of the experiment. Allowable values = 30-45 for dung and NA for mopane. Units = g. Dung was weighed within a litterbag while mopane was weighed in a weigh boat and then transferred to a litterbag.

`InitialWetMass`: The initial wet mass of the litter at the start of the experiment. Allowable values = 27.5-41 for dung and 3-3.5 for mopane. This is a calculated varibale for dung and a measure variable for mopane. Units = g.

`DryAndBoat`: The mass of the dried litter and the weigh boat after drying at the end of the experiment. Allowable values = 0-50 for dung and 0-20 for mopane. Units = g.

`Boat`: The mass of the weigh boat used to hold the mopane or dung litter during drying. Allowable values = 0-40. 0 indicates that the weigh boat was tared prior to weighing (boats were typically tared when there was only one balance user working at a time). Units = g.

`BagMass`: The mass of the empty bag used to hold the litter during wet-dry correction. Allowable values = 3 - 3.5. Units = g.

`WetMassAndBag`: The mass of the wet litter and the bag used to hold the litter during wet-dry correction. Allowable values = 5 - 50. Units = g.

`DryMassAndBag`: The mass of the dried litter and the bag used to hold the litter during wet-dry correction. Allowable values = 6 - 22. Units = g.

`CrucibleLid`: The mass of the crucible and lid before adding litter. Allowable values = 15 - 20 for both dung and mopane. Units = g.

`CrucibldLidLitter`: The mass of the crucible, lid, and dried litter before ashing. Allowable values = 18 - 30. Units = g.

`CrucibleLidAsh`: The mass of the crucible, lid, and ash after ashing. Allowable values = 18 - 30. Units = g.

#### Calculated variables

`WetDryCorrection`: The wet-dry correction factor calculated as the proportion of dry mass to wet mass for each litter type. This factor is used to convert initial wet mass to initial dry mass. Allowable values = 0.379 for dung and 0.907 for mopane. Unitless.

`InitialDryMass`: The initial dry mass of the litter at the start of the experiment (after wet-dry correction). InitialDryMass = InitialWetMass x ProportionDry. Allowable values = 10 - 16 for dung and 2.6 - 3.1 for mopane. Units = g.

`FinalDryMass`: The final dry mass of the litter at the end of the experiment (after wet-dry correction). FinalDryMass = DryAndBoat - Boat. Allowable values = 0 - 65. Units = g.

`TransportLossCorrection`: The transport loss correction factor calculated as the mean proportion of final dry mass to initial dry mass for time 0 litterbags for each litter type. This factor is used to correct initial dry mass for transport losses. Allowable values = 0.948 for dung and 0.998 for mopane. Unitless.

`InitialDryMass_TransportCorrected`: The initial dry mass of the litter at the start of the experiment after correcting for transport losses (mass loss in time 0 bags). InitialDryMass_TransportCorrected = InitialDryMass x TransportLossCorrection. Allowable values = 10 - 16 for dung and 2.6 - 3.1 for mopane. Units = g.

`ProportionAsh`: The proportion of ash in the litter calculated as the mass of ash divided by the mass of litter for each litterbag after collection. Allowable values = 0.01 - 1. Unitless.

`AshCorrection`: The ash correction factor calculated as the mean proportion of ash in time 0 litterbags for each litter type. This factor is used to correct initial dry mass for ash content. Allowable values = 0.088 for dung and 0.064 for mopane. Unitless.

`Initial_AFDM_TransCorrected`: The ash-free dry mass of the litter at the start of the experiment after correcting for transport losses and ash content. Initial_AFDM_TransCorrected = InitialDryMass_TransportCorrected x (1 - AshCorrection). Allowable values = 10 - 16 for dung and 2.6 - 3.1 for mopane. Units = g.

`Final_AFDM`: The ash-free dry mass of the litter at the end of the experiment after correcting for ash content. Final_AFDM = FinalDryMass x (1 - ProportionAsh). Allowable values = 0 - 16 for dung and 0 - 3.1 for mopane. Units = g.

`PctAFDM_Remaining`: The percentage of ash-free dry mass remaining at the end of the experiment. PctAFDM_Remaining = (Final_AFDM / Initial_AFDM_TransCorrected) \* 100. Allowable values = 0 - 100 for both dung and mopane. Units = %.

`WetMass`: The wet mass of the litter used in the wet-dry correction. WetMass = WetMassAndBag - BagMass. Allowable values = 3 - 50. Units = g.

`DryMass`: The dry mass of the litter used in the wet-dry correction. DryMass = DryMassAndBag - BagMass. Allowable values = 3 - 18. Units = g.

`ProportionDry`: The proportion of dry mass to wet mass for each litterbag used in the wet-dry correction. ProportionDry = DryMass / WetMass. Allowable values = 0.1 - 1. Unitless.

`LitterMass`: The mass of the litter used in the ash correction. LitterMass = CrucibleLidLitter - CrucibleLid. Allowable values = 3 - 18. Units = g.

`AshMass`: The mass of the ash remaining after ashing the litter. AshMass = CrucibleLidAsh - CrucibleLid. Allowable values = 0.01 - 5. Units = g.

## Preparation Steps

```{r}
#| label: load-packages

library(tidyverse)
# add other packages here (note that any "install.packages" calls should be made in the console and not on the script as it isn't nice to inadvertently install something on someone else's computer.)

```

```{r}
#| label: set-up-read-write-filepath

# This call uses the stored config file to define the path on your computer for the DERT Data & Metadata Dropbox folder. Find detailed instructions on this set-up in "Field and Lab SOPS/Data SOPs/DERT Data and Code Storage.docx"
DERT_Archives <- config::get()
```

```{r}
#| label: read-in-data

# Read in your data files (this uses an example from the DERT Dropbox - replace with your file)
# Be sure to describe what data you are pulling in. 
swc <- read_csv(file.path(DERT_Archives, "Amari Compost Incubations", 
              "FinalData", "SWC_With_Compost_HT.csv"),
                show_col_types = FALSE)
```

### Data Cleaning and Work Up

```{r}
#| label: data-cleaning

# likely there is a lot of data cleaning that you'll do, so you will likely have multiple code chunks of data cleaning. Add clear documentation for decisions (e.g., it is fine to discard data that you have a clear rationale that are suspect. Here is a great place to do it, just document it carefully.)

```

```{r}
#| label: data-work-up

# Here is a good time to make sure that your data type is defined properly and that if you have categorical data your levels are set properly. In this case, we're making sure that the treatments are factors and making them appear in a logical (not alphabetical) order.

# define compost and water treatments as factors
combined_df$CompTreat<-factor(combined_df$CompTreat,
                               levels = c("Blank","S","C15","C30"))
combined_df$WaterTreat<-factor(combined_df$WaterTreat,
                            levels = c("Blank","D", "W05","W10","W15"))


```

###Visualization and Save Data

```{r}
#| label: Exploratory-visualization

# probably you want to do some histograms or other such explorations

```

```{r}
#| label: export-data

# you may want to save your data at this point. Maybe you brought in L1 data and now it is L2? Here, we are saving the "df_L2" to the L2 folder in Dropbox. Make sure that your Dropbox folder has the appropriate folders created before running this code.

write_csv(
  dataframe_L2,
  file = file.path(
    DERT_Archives,
    "Data",
    "L2",
    "MyAwesomeData_L2.csv"
  )
)

```

## Data Analysis

### Regressions

```{r}
#| label: swc-regression

# or whatever analyses are you doing...

```

### ANOVAs

```{r}
#| label: ANOVA-with-saved-outputs

# This is some example code for an ANOVA. Note that the data are written as a text file to Dropbox after the analysis. Make sure that you have the appropriate folders created in Dropbox before running this code.

# create a df of mean swc values for each jar
swc_jar_means_df <- combined_df |>
  filter(WaterTreat != "Blank") |>
  filter(WaterTreat != "D") |>
  filter(floor(Day) == Day) |> # remove non-integer values for day (duplicate measurements)
  group_by(Jar, CompTreat, WaterTreat) |>
  summarise(
    N = sum(!is.na(SWC)),
    mean_SWC = mean(SWC, na.rm=TRUE),
    sd_SWC   = sd(SWC, na.rm=TRUE),
    se_SWC   = sd_SWC / sqrt(N)
               ) 

#ANOVA test
swc.anova <- aov(mean_SWC ~ CompTreat*WaterTreat, data = swc_jar_means_df)
summary(swc.anova)
# Water and compost and W*C are all significant

#Post-hoc test
TukeyHSD(swc.anova)
#A few non-significant groups

#Checking assumptions
#Summary statistics
model.tables(swc.anova, type="means", se = TRUE)

#Checking for equal variance
plot(swc.anova, 1)
leveneTest(mean_SWC ~ CompTreat*WaterTreat, data = swc_jar_means_df)
#Does have equal variance

#Checking for normal distribution
plot(swc.anova, 2)
anova_residuals <- residuals(object = swc.anova)
shapiro.test(x = anova_residuals)
#Is normally distributed

## Save results to a text file

# Set output file
out_file <- file.path(DERT_Archives, "YourFolderHere", "Results", "swc_anova_results.txt")

# Write results to file
sink(out_file)

cat("=== ANOVA Summary ===\n")
print(summary(swc.anova))

cat("\n=== TukeyHSD Post-hoc ===\n")
print(TukeyHSD(swc.anova))

cat("\n=== Model Tables (means & SE) ===\n")
print(model.tables(swc.anova, type = "means", se = TRUE))

cat("\n=== Levene's Test for Homogeneity of Variance ===\n")
print(leveneTest(mean_SWC ~ CompTreat * WaterTreat, data = swc_jar_means_df))

cat("\n=== Shapiro-Wilk Test for Normality of Residuals ===\n")
anova_residuals <- residuals(object = swc.anova)
print(shapiro.test(x = anova_residuals))

sink()  # close connection
```

## Figures

```{r}
#label: soil-water-content-bar-graph

# Here's a random bar chart example. Note at the bottom that the figure is saved to Dropbox. Make sure that you have the appropriate folders created in Dropbox before running this code.

plot2 <- Jar_CN_means_incub.only |>
  ggplot(aes(x=CompTreat, y=Jar_mean_N, fill=WaterTreat)) + 
  geom_boxplot() +
  labs(x="Compost Treatment", 
        y = "Soil Nitrogen (% by mass)") +
  theme_bw() +
  # Set colors, order, and labels for WaterTreat levels
  scale_fill_manual(values = c("blue", "darkgray", "red"), 
                     breaks = c("W15", "W10", "W05"),
                     labels = c("large-infrequent", "intermediate", "small-frequent")) +  
  labs(fill = NULL) + # remove legend title
  scale_x_discrete(breaks = c("S", "C15", "C30"),
      labels = c("No Compost", "Low Compost", "High Compost")) +
  theme(legend.position = c(0.85, 0.3),  # Adjust legend position as needed
        legend.justification = c(1, 1),
        text = element_text(size = 18),  # Increase all font sizes
        axis.text = element_text(size = 16),  # Increase axis text size
        axis.title = element_text(size = 18),  # Increase axis title size
        legend.text = element_text(size = 16),  # Increase legend text size
       panel.grid = element_blank()) +  # Remove gridlines
       geom_hline(yintercept=pctN.origsoil, linetype="dashed", 
                color = "black", linewidth=1)
plot2

# Export plot
ggsave(
  filename = file.path(DERT_Archives, "YourFolderHere", 
                       "Figures", "pctN_plot.jpg"),
  plot = plot2,
  width = 7, height = 5, dpi = 300
)
```
